\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-04-27 forwards-and-backwards.tex
put back in general kernel for PME

\begin{proposition}
	\label{nonlocalPME to PME}
Suppose that
% we can write $\rho_\gamma=\zeta*\check{\zeta}$, where $\check{\zeta}(x)=\zeta(-x)$ and
% $\zeta\in\mathcal{S}(\IR^d)$ (the Schwartz space of rapidly decreasing functions).
$\rho_\gamma = p_{\epsilon^2_\gamma}$,
and that %$\psi_0^\epsilon\in L^1(\IR^d)$, 
$\psi_0^\epsilon\geq 0$ is such that
there exist $\lambda\in (0,1)$ and $C\in (0,\infty)$ (independent
of $\epsilon$) such that 
\begin{equation*}
\int\exp(\lambda \|x\|)\psi_0^\epsilon(x)dx<C,\quad\mbox{ and }
\sup_\epsilon\int \psi_0^\epsilon|\log \psi_0^\epsilon|dx<\infty,
\end{equation*}
with
$\psi_0^\epsilon\rightharpoonup \psi_0$ as $\epsilon\to 0$. Then
writing $\psi^\epsilon$ for the solution to~(\ref{mollified equation})
on $[0,T]\times \IR^d$ with
initial condition $\psi_0^\epsilon$,
$\psi^\epsilon\rightharpoonup \psi$ as $\epsilon\to 0$ where
$\psi\in L_t^2(H^1)\cap C_t(L^1)$, $\int \psi|\log \psi| dx<\infty$, and
$\psi$ solves~(\ref{PME}) on $[0,T]\times \IR^d$. 
\end{proposition}
The example that we have in mind for the kernel $\rho_\gamma$ is a Gaussian
kernel.
\begin{remark}
Although it seems hard to formulate an all-encompassing result,
Propositions~\ref{prop:nonlocal_to_local} and~\ref{nonlocalPME to PME} are by no
means exhaustive. When the scaling limit is deterministic, one can 
expect analogous results under rather general conditions. However, when the limit
points are stochastic, they resemble ``nonlinear superprocesses'' and so one cannot
expect a density with respect to Lebesgue measure in $d\geq 2$. It is then not
reasonable to 
expect to be able to make sense of the limit if we scale the kernels in this way.
Moreover, in one dimension, where the classical superprocess does have a density with 
respect to Lebesgue measure, the form of~(\ref{eqn:limiting_mgale_problem})
suggests that even if one can remove the local averaging from $\gamma$, it
will be necessary
to retain averaging of $r$ in order to obtain a well-defined limit. 
\end{remark}

@@

In this subsection we prove Proposition~\ref{nonlocalPME to PME}.
To ease notation, we present
the proof in $d=1$ (although we retain the notation $\nabla$).
However to recall the dependence on $\epsilon$ we write $\rho^\epsilon$ for $\rho_\gamma$
(which recall for this Proposition we suppose is Gaussian
and equal to $p_{\epsilon^2_\gamma}$).
It should be clear that it extends almost without change
to higher dimensions.

Recall that 
we are concerned with non-negative solutions to the equation~(\ref{mollified equation}):
\begin{equation*}
	\partial_t\psi_t^\epsilon(x) =
	\Delta\left(\psi_t^\epsilon \, \rho^\epsilon *\psi_t^\epsilon\right)(x)
	+\psi_t^\epsilon(x)\left(1-\rho^\epsilon *\psi_t^\epsilon(x) 
\right) .
\end{equation*}
% and we assume that $\rho=\zeta*\check{\zeta}$ with $\zeta$ a rapidly decreasing function. 
% The example we have in mind is $\zeta$ (and therefore $\rho$) being
We shall prove that under the 
assumptions of Proposition~\ref{nonlocalPME to PME},
as $\epsilon\to 0$, we have
convergence to the solution to the porous medium equation with
logistic growth,
equation~\eqref{PME1}:
\begin{equation*}
	\partial_t \psi_t(x)=
	\Delta\left(\psi_t^2\right)(x)
	+\psi_t(x)\left(1-\psi_t(x)\right).
\end{equation*}

@@

\begin{proof}[Proof of Proposition~\ref{nonlocalPME to PME}]
Write $\zeta = p_{\epsilon^2_\gamma / 2}$, so that $\rho = \zeta * \zeta$.
First observe that
\begin{multline*}
	\int \psi_t^\epsilon(x)\, \rho^{\epsilon}*\psi_t^\epsilon(x) dx 
=
\int\int\int \psi_t^\epsilon(x)\psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
{\zeta}^{\epsilon}(z) dz dy dx 
\\
=
\int\int\int \psi_t^\epsilon(\tilde{x}-\tilde{z})
\psi_t^\epsilon(\tilde{x}-\tilde{y})
\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z}) d\tilde{z} d\tilde{y} d \tilde{x}
	=\int\left(\zeta^{\epsilon}*\psi_t^\epsilon(x)\right)^2 dx,
\end{multline*}
where we have set $\tilde{x}=x-z$, $\tilde{y}=y-z$, $\tilde{z}=-z$.

@@

We shall next show that $\zeta^\epsilon * \psi_t^\epsilon$ solves \eqref{PME1}
up to a remainder of order $\epsilon$.
First observe that
\begin{equation}
\label{total equation}
\int\Delta\left((\rho^{\epsilon}*\psi_t^\epsilon)\psi_t^\epsilon\right)\phi  dx
    =
    - \int \nabla(\rho^{\epsilon}*\psi_t^\epsilon)\, \psi_t^\epsilon\,\nabla\phi dx
    -\int \rho^{\epsilon}*\psi_t^\epsilon \,\nabla \psi_t^\epsilon \, \nabla\phi dx.
\end{equation}
We would like to show that this is close to $\int (\zeta^\epsilon * \psi^\epsilon_t)^2 \Delta \phi dx$.
For the first term
\begin{eqnarray}
\nonumber
\int \nabla(\rho^{\epsilon}*\psi_t^\epsilon)\, \psi_t^\epsilon\, \nabla\phi  dx
&=&
\int\int\int\nabla \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
{\zeta}^{\epsilon}(z)
\psi_t^\epsilon(x)\nabla\phi(x) dz  dy  dx
\\
\nonumber
&=&
\int\int\int\nabla \psi_t^\epsilon(\tilde{x}-\tilde{y})
\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z})\psi_t^\epsilon(\tilde{x}-\tilde{z})
\nabla\phi(\tilde{x}-\tilde{z}) d\tilde{z} d\tilde{y} d\tilde{x}
\\
\nonumber
&=&\int(\nabla \zeta^{\epsilon}*\psi_t^\epsilon)\,
\left(\zeta^{\epsilon}*(\psi_t^\epsilon\nabla\phi)\right)
d x\\
\nonumber
&=&
\frac{1}{2}\int\nabla ((\zeta^{\epsilon}*\psi_t^\epsilon)^2)
\,\nabla\phi  dx
\\
\label{error 1}
&&+
\int\nabla(\zeta^{\epsilon}*\psi_t^\epsilon)
\left[\zeta^{\epsilon}*(\psi_t^\epsilon\,\nabla\phi)-
\nabla\phi\,(\zeta^{\epsilon}*\psi_t^\epsilon)\right] dx,
\end{eqnarray}
where, as before, we have substituted $\tilde{x}=x-z$, $\tilde{y}=y-z$,
$\tilde{z}=-z$.
To control the term~(\ref{error 1})
we use the Intermediate Value Theorem to see that
\begin{multline*}
\left|\int\left[\psi_t^\epsilon(x-y)\nabla\phi(x-y)\zeta^{\epsilon}(y)-
\nabla\phi(x)\psi_t^\epsilon(x-y)\zeta^{\epsilon}(y)
\right] dy\right|
\\
\leq C\|\Delta\phi\|_\infty\int \psi_t^\epsilon (x-y)y\zeta^{\epsilon}(y) d y.
\end{multline*}
Since $\zeta\in\mathcal{S}(\IR)$, 
the integral in this expression is
$\mathcal{O}(\epsilon)$.

Similarly, for the second term in~(\ref{total equation}),
\begin{eqnarray}
\nonumber
\int (\rho^{\epsilon}*\psi_t^\epsilon)\, \nabla \psi_t^\epsilon\, \nabla\phi dx
&=&
\int\int\int \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
{\zeta}^{\epsilon}(z)\nabla \psi_t^\epsilon(x)\nabla\phi(x) dz dy  dx
\\
\nonumber
&=&
\int\int\int \psi_t^\epsilon(\tilde{x}-\tilde{y})\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z})\nabla \psi_t^\epsilon(\tilde{x}-\tilde{z})
\nabla\phi(\tilde{x}-\tilde{z}) d \tilde{z} d\tilde{y} d\tilde{x}
\\
\nonumber
	&=&\int(\zeta^{\epsilon}*\psi_t^\epsilon)\,
\left(\zeta^{\epsilon}*(\nabla \psi_t^\epsilon \, \nabla\phi)\right)  dx\\
\nonumber
&=&
\frac{1}{2}\int\nabla ((\zeta^{\epsilon}*\psi_t^\epsilon)^2)
\, \nabla\phi  dx
\\
\label{error 2}
&&+
\int (\zeta^{\epsilon}*\psi_t^\epsilon)
\left[\zeta^{\epsilon}*(\nabla \psi_t^\epsilon\, \nabla\phi)-
\nabla\phi\,(\nabla \zeta^{\epsilon}*\psi_t^\epsilon)\right] dx,
\end{eqnarray}
and~(\ref{error 2}) is controlled in the same way as~(\ref{error 1}):
using the Intermediate Value Theorem,
\begin{multline*}
\left|\int\left[\nabla \psi_t^\epsilon(x-y)\nabla\phi(x-y)\zeta^{\epsilon}(y)
-\nabla\phi(x)\nabla \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y)
\right] dy\right|
\\
\leq C\|\Delta\phi\|_\infty\left|\int \nabla \psi_t^\epsilon(x-y)
y \zeta^{\epsilon}(y)
d y\right|,
\end{multline*}
which again is $\mathcal{O}(\epsilon)$.

>>>>

\begin{proposition}
	\label{nonlocalPME to PME}
Suppose that
we can write $\rho_\gamma=\zeta*\check{\zeta}$, where $\check{\zeta}(x)=\zeta(-x)$ and
$\zeta\in\mathcal{S}(\IR^d)$ (the Schwartz space of rapidly decreasing functions).
Furthermore, suppose that %$\psi_0^\epsilon\in L^1(\IR^d)$, 
$\psi_0^\epsilon\geq 0$ is such that
there exist $\lambda\in (0,1)$ and $C\in (0,\infty)$ (independent
of $\epsilon$) such that 
\begin{equation*}
\int\exp(\lambda \|x\|)\psi_0^\epsilon(x)dx<C,\quad\mbox{ and }
\sup_\epsilon\int \psi_0^\epsilon|\log \psi_0^\epsilon|dx<\infty,
\end{equation*}
with
$\psi_0^\epsilon\rightharpoonup \psi_0$ as $\epsilon\to 0$. Then
writing $\psi^\epsilon$ for the solution to~(\ref{mollified equation})
on $[0,T]\times \IR^d$ with
initial condition $\psi_0^\epsilon$,
$\psi^\epsilon\rightharpoonup \psi$ as $\epsilon\to 0$ where
$\psi\in L_t^2(H^1)\cap C_t(L^1)$, $\int \psi|\log \psi| dx<\infty$, and
$\psi$ solves~(\ref{PME}) on $[0,T]\times \IR^d$. 
\end{proposition}
The example that we have in mind for the kernel $\rho_\gamma$ is a Gaussian
kernel.
\begin{remark}
Although it seems hard to formulate an all-encompassing result,
Propositions~\ref{prop:nonlocal_to_local} and~\ref{nonlocalPME to PME} are by no
means exhaustive. When the scaling limit is deterministic, one can 
expect analogous results under rather general conditions. However, when the limit
points are stochastic, they resemble ``nonlinear superprocesses'' and so one cannot
expect a density with respect to Lebesgue measure in $d\geq 2$. It is then not
reasonable to 
expect to be able to make sense of the limit if we scale the kernels in this way.
Moreover, in one dimension, where the classical superprocess does have a density with 
respect to Lebesgue measure, the form of~(\ref{eqn:limiting_mgale_problem})
suggests that even if one can remove the local averaging from $\gamma$, it
will be necessary
to retain averaging of $r$ in order to obtain a well-defined limit. 
\end{remark}

@@

In this subsection we prove Proposition~\ref{nonlocalPME to PME}.
To ease notation, we present
the proof in $d=1$ (although we retain the notation $\nabla$).
However, to recall the dependence on $\epsilon$ we write $\rho^\epsilon$ for $\rho_\gamma$.
It should be clear that it extends almost without change
to higher dimensions.

Recall that 
we are concerned with non-negative solutions to the equation~(\ref{mollified equation}):
\begin{equation*}
	\partial_t\psi_t^\epsilon(x) =
	\Delta\left(\psi_t^\epsilon \, \rho^\epsilon *\psi_t^\epsilon\right)(x)
	+\psi_t^\epsilon(x)\left(1-\rho^\epsilon *\psi_t^\epsilon(x) 
\right) .
\end{equation*}
and 
we assume that $\rho=\zeta*\check{\zeta}$ with $\zeta$ a rapidly decreasing function. 
The example we have in mind is $\zeta$ (and therefore $\rho$) being the density of
a mean zero Gaussian random variable.  
We shall prove that under the 
assumptions of Proposition~\ref{nonlocalPME to PME},
as $\epsilon\to 0$, we have
convergence to the solution to the porous medium equation with
logistic growth,
equation~\eqref{PME1}:
\begin{equation*}
	\partial_t \psi_t(x)=
	\Delta\left(\psi_t^2\right)(x)
	+\psi_t(x)\left(1-\psi_t(x)\right).
\end{equation*}

@@

\begin{proof}[Proof of Proposition~\ref{nonlocalPME to PME}]
First observe that
\begin{multline*}
	\int \psi_t^\epsilon(x)\, \rho^{\epsilon}*\psi_t^\epsilon(x) dx 
=
\int\int\int \psi_t^\epsilon(x)\psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
\check{\zeta}^{\epsilon}(z) dz dy dx 
\\
=
\int\int\int \psi_t^\epsilon(\tilde{x}-\tilde{z})
\psi_t^\epsilon(\tilde{x}-\tilde{y})
\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z}) d\tilde{z} d\tilde{y} d \tilde{x}
	=\int\left(\zeta^{\epsilon}*\psi_t^\epsilon(x)\right)^2 dx,
\end{multline*}
where we have set $\tilde{x}=x-z$, $\tilde{y}=y-z$, $\tilde{z}=-z$.

@@

We shall next show that $\zeta^\epsilon * \psi_t^\epsilon$ solves \eqref{PME1}
up to a remainder of order $\epsilon$.
First observe that
\begin{equation}
\label{total equation}
\int\Delta\left((\rho^{\epsilon}*\psi_t^\epsilon)\psi_t^\epsilon\right)\phi  dx
    =
    - \int \nabla(\rho^{\epsilon}*\psi_t^\epsilon)\, \psi_t^\epsilon\,\nabla\phi dx
    -\int \rho^{\epsilon}*\psi_t^\epsilon \,\nabla \psi_t^\epsilon \, \nabla\phi dx.
\end{equation}
We would like to show that this is close to $\int (\zeta^\epsilon * \psi^\epsilon_t)^2 \Delta \phi dx$.
For the first term
\begin{eqnarray}
\nonumber
\int \nabla(\rho^{\epsilon}*\psi_t^\epsilon)\, \psi_t^\epsilon\, \nabla\phi  dx
&=&
\int\int\int\nabla \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
\check{\zeta}^{\epsilon}(z)
\psi_t^\epsilon(x)\nabla\phi(x) dz  dy  dx
\\
\nonumber
&=&
\int\int\int\nabla \psi_t^\epsilon(\tilde{x}-\tilde{y})
\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z})\psi_t^\epsilon(\tilde{x}-\tilde{z})
\nabla\phi(\tilde{x}-\tilde{z}) d\tilde{z} d\tilde{y} d\tilde{x}
\\
\nonumber
&=&\int(\nabla \zeta^{\epsilon}*\psi_t^\epsilon)\,
\left(\zeta^{\epsilon}*(\psi_t^\epsilon\nabla\phi)\right)
d x\\
\nonumber
&=&
\frac{1}{2}\int\nabla ((\zeta^{\epsilon}*\psi_t^\epsilon)^2)
\,\nabla\phi  dx
\\
\label{error 1}
&&+
\int\nabla(\zeta^{\epsilon}*\psi_t^\epsilon)
\left[\zeta^{\epsilon}*(\psi_t^\epsilon\,\nabla\phi)-
\nabla\phi\,(\zeta^{\epsilon}*\psi_t^\epsilon)\right] dx,
\end{eqnarray}
where, as before, we have substituted $\tilde{x}=x-z$, $\tilde{y}=y-z$,
$\tilde{z}=-z$.
To control the term~(\ref{error 1})
we use the Intermediate Value Theorem to see that
\begin{multline*}
\left|\int\left[\psi_t^\epsilon(x-y)\nabla\phi(x-y)\zeta^{\epsilon}(y)-
\nabla\phi(x)\psi_t^\epsilon(x-y)\zeta^{\epsilon}(y)
\right] dy\right|
\\
\leq C\|\Delta\phi\|_\infty\int \psi_t^\epsilon (x-y)y\zeta^{\epsilon}(y) d y.
\end{multline*}
Since $\zeta\in\mathcal{S}(\IR)$, 
the integral in this expression is
$\mathcal{O}(\epsilon)$.

Similarly, for the second term in~(\ref{total equation}),
\begin{eqnarray}
\nonumber
\int (\rho^{\epsilon}*\psi_t^\epsilon)\, \nabla \psi_t^\epsilon\, \nabla\phi dx
&=&
\int\int\int \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
\check{\zeta}^{\epsilon}(z)\nabla \psi_t^\epsilon(x)\nabla\phi(x) dz dy  dx
\\
\nonumber
&=&
\int\int\int \psi_t^\epsilon(\tilde{x}-\tilde{y})\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z})\nabla \psi_t^\epsilon(\tilde{x}-\tilde{z})
\nabla\phi(\tilde{x}-\tilde{z}) d \tilde{z} d\tilde{y} d\tilde{x}
\\
\nonumber
	&=&\int(\zeta^{\epsilon}*\psi_t^\epsilon)\,
\left(\zeta^{\epsilon}*(\nabla \psi_t^\epsilon \, \nabla\phi)\right)  dx\\
\nonumber
&=&
\frac{1}{2}\int\nabla ((\zeta^{\epsilon}*\psi_t^\epsilon)^2)
\, \nabla\phi  dx
\\
\label{error 2}
&&+
\int (\zeta^{\epsilon}*\psi_t^\epsilon)
\left[\zeta^{\epsilon}*(\nabla \psi_t^\epsilon\, \nabla\phi)-
\nabla\phi\,(\nabla \zeta^{\epsilon}*\psi_t^\epsilon)\right] dx,
\end{eqnarray}
and~(\ref{error 2}) is controlled in the same way as~(\ref{error 1}):
using the Intermediate Value Theorem,
\begin{multline*}
\left|\int\left[\nabla \psi_t^\epsilon(x-y)\nabla\phi(x-y)\zeta^{\epsilon}(y)
-\nabla\phi(x)\nabla \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y)
\right] dy\right|
\\
\leq C\|\Delta\phi\|_\infty\left|\int \nabla \psi_t^\epsilon(x-y)
y \zeta^{\epsilon}(y)
d y\right|,
\end{multline*}
which again is $\mathcal{O}(\epsilon)$.
